Assessing methods to adjust for assay differences across COVID-19 serological surveillance studies in Canada

### 2024-01-19 5_manuscript/manuscript.qmd
Minor updates on methods, add topics to discussion

-   APL assay data before the Omicron wave was mostly missing, Carmen said they have already approved the data access but the data is still on its way.

-   In AB Outpatient Lab Data, they changed the assay (Diasorin S --\> Abbott S) in the midway, this would affect the validity of analysis based on quantitative test results, as the results from different assay are not directly comparable. Hopefully this can be clarified in the updated version of the dataset.

@@

@@

\[insert @thm-data here if necessary\]

@@

$$

$$

Bayesian models adjusting for waning immunity:
$$

$$

## **Proportion sufficiently protected:**

@@

\[Insert @fig-immunity here if necessary\]

@@

In our case, the sensitivity is where most variance come from. Specificity is good over time, and we can used the number published by the manufacturer (Roche & Abbott). Since sens is imperfect (0.8--1.0) and specificity is mostly 100% trustworthy (\~1.0), the R-G equation and be simplified as:

@@

Between data (APL, CBS, CANPATH) diff is not the major focus of this study, Yuan's paper will focus on those. For this paper, the major focus should be the difference between adjusting methods

@@

>>>>

-   APL assay data before the Omicron wave was mostly missing, Carmen said they have already approved the data access but the data is still on its way. Updates Jan-16-2024: Carmen said more data will be available soon.

-   In AB Outpatient Lab Data, they changed the assay (Diasorin S --\> Abbott S) in the midway, this would affect the validity of analysis based on quantitative test results, as the results from different assay are not directly comparable. Hopefully this can be clarified in the updated version of the dataset. Our currently analysis assume the Anti-S used in omicron wave are all Abbott's, based on the paper published by Carmen in the post-vax stage: https://doi.org/10.1080/23744235.2022.2080250

@@

**Notes**: *The paragraph below describes briefly about the time frame of data collection in those datasets. CBS is still actively updating (last ver downloaded on Nov 30th) and APL is also actively updating, we are currently waiting for additional linked variables from Carmen.*

@@

[insert @thm-data here if necessary]

@@

\$\$

\$\$

Bayesian models adjusting for waning immunity: \$\$

\$\$

## **NT50 and corresponding antibody level:**

@@

[Insert @fig-immunity here if necessary]

@@

In our case, the sensitivity is where most variance come from. Specificity is good over time, and we shall not only take the numbers published by the manufacturer (Roche & Abbott) into consideration. Since sens is imperfect (0.8--1.0) and specificity is mostly 100% trustworthy (\~1.0), the R-G equation can be simplified as:

@@

**Notes**: *Variances among data (APL, CBS, CANPATH) is not the major focus of this study, Yuan's paper will focus on those by evaluating the representativeness of the serosurveillance data. For this paper, the major focus should be the difference between adjusting methods. The potential readers of this papers includes those who are using serosurveillance data to monitor the progress of Covid pandemic, or other pathogens of interests which possess similar features as SARS-CoV-2.*

The take-home messages would be:

1.  Without any adjustment, the public health inference on test positivity is very limited. It only estimates the proportion of test positive results among blood samples collected in the serosurveillance study.
2.  Adjusting for certain factors will change the public health inference, meanwhile it will also improve its generalizability. When the imperfect nature of the seroassays is taken into account, the adjusted proportion of positive test results (seroprevalence) will be able to represent the target population in our target population, from which our sample has been collected. Upon the seroprevalence, after adjusting for seroreversion using Bayesian models, we get attack rates of SARS-CoV-2, which estimates the proportion of population infected by the virus since the beginning of pandemic.
3.  There's gap among all those different measurements, and also they have different public health inference. Future researchers should be cautious when making references to those different measurements.

@@

## Supplement Methods

**Notes**: *Current ver of APL has completed very limited data linkage, which only includes PCR test results/time and vaccination time. The paragraph briefly describe the data linkage for three datasets, however for now we do not need those variables from linked administrative data.*
